Fusing Filters with Integer Linear Programming

### 2014-04-20 2014betterfusionforfilters/section/03-SizeInference.tex
minor size inf
constraint generation fixes
change constraint generation to array, somewhat prettier
more size inference

Size inference has been explored before in the context of fusion by Chatterjee~\cite{chatterjee1991size}, but the only constraints they support are equality.
Their formulation has no filtering operations, and all array operations in their case generate the same size as at least one of the operation's inputs.
A more in-depth analysis has also been explored by Jay~\cite{jay1996shape} in the form of shape inference, which is built from primitives such as @cons@ and @nil@.
Since we have fewer and simpler primitives, our size inference does not need to be as complicated as Jay's shape inference.

\begin{tabbing}
MMMMM       \= MM \= MMMMMMMMMMM \= \kill
$rate$      \> @::=@ \> $k_n$               \> (rate variable)\\
            \> $~|$  \> $rate \times rate$  \> (cross product) \\
%            \> $~|$  \> $Filtered_n$        \> (output of @filter@s) \\
%            \> $~|$  \> $External_n$        \> (output of @external@s) \\
\\
$constraint$ \> @::=@ \> $rate = rate$                          \> (equality constraint) \\
             \> $~|$  \> $constraint \wedge constraint$         \> (conjunction) \\
             \> $~|$  \> $\forall k_n.\ constraint$              \> (forall quantification) \\
             \> $~|$  \> $\exists k_n.\ constraint$              \> (exists with no bounds)\\
             \> $~|$  \> $\exists k_n \le rate.\ constraint$     \> (exists with upper bound)    \\
\\
$\tau$       \> @::=@ \> $rate$                                  \> (loop size of some rate) \\
             \> $~|$  \> @external@                              \> (external and unfusible) \\
\end{tabbing}

\newcommand{\constraint}[3]{#1 \vDash #2 : #3}
\newcommand{\inferC}[2]{
    $$\begin{array}{c}
        \infer{#1}{#2}
    \end{array}$$}
\newcommand{\ruleIC}[3]{
    \inferC
        {\constraint{#1; binds}{#2 C}{#3 ~ T}}
        {\constraint{binds}{C}{T}}}

\[ \constraint{binds}{constraint}{(\tau_{name} \Rightarrow \tau)} \]

@@

\vspace{1em}

\inferC{\constraint{\lambda scalars~arrays \to @let @ binds}{\forall_{a \in arrays} k_a. C}{T}}
       {\constraint{binds}{C}{T}}

\vspace{1em}

\ruleIC{o = @fold@~f~n}{}{\tau_o = k_n, }

\vspace{1em}

\ruleIC{o = @map@_n~f~ns}{\forall k_o.\ \bigwedge_{n \in ns}\{k_o = k_n\} \wedge}{\tau_o = k_o, }

\vspace{1em}

\ruleIC{o = @filter@~f~n}{\exists k_o \le k_n.\ }{\tau_o = k_n, }

@@

\vspace{1em}

\ruleIC{o = @gather@~i~d}{\forall k_o.\ k_o = k_i \wedge}{\tau_o = k_i, }

\vspace{1em}

\ruleIC{o = @cross@~a~b}{\forall k_o.\ k_o = k_a \times k_b \wedge C}{\tau_o = k_a \times k_b, }

\vspace{1em}

\ruleIC{outs = @external@~ins}{\exists_{o \in outs} k_o.\ }{\tau_{outs} = @external@, }

\vspace{1em}

>>>>

Size inference has been explored before in the context of fusion by Chatterjee~\cite{chatterjee1991size}, but the only constraints they support are equality.
Their formulation has no filtering operations, and all array operations in their case generate the same size as at least one of the operation's inputs.
A more in-depth analysis has also been explored by Jay~\cite{jay1996shape} in the form of shape inference, which is built from primitives such as @cons@ and @nil@.
Since we have fewer primitives and simpler goals, our size inference does not need to be as complicated as Jay's shape inference.

\newcommand{\constr}[1]{\llbracket #1 \rrbracket}
\newcommand{\hole}[0]{[]}
\newcommand{\fillhole}[2]{#1\Left[#2\right]}

\begin{tabbing}
MMMMM       \= MM \= MMMMMMMMMMM \= \kill
$rate$      \> @::=@ \> $k_n$               \> (rate variable)\\
            \> $~|$  \> $rate \times rate$  \> (cross product) \\
%            \> $~|$  \> $Filtered_n$        \> (output of @filter@s) \\
%            \> $~|$  \> $External_n$        \> (output of @external@s) \\
\\
$C$          \> @::=@ \> $true$                                 \> (trivially true) \\
             \> $~|$  \> $\hole$                                \> (constraint hole) \\
             \> $~|$  \> $rate = rate$                          \> (equality constraint) \\
             \> $~|$  \> $C \wedge C$                           \> (conjunction) \\
             \> $~|$  \> $\forall k_n.\ C$                      \> (forall quantification) \\
             \> $~|$  \> $\exists k_n.\ C$                      \> (exists with no bounds)\\
             \> $~|$  \> $\exists k_n \le rate.\ C$             \> (exists with upper bound)    \\
\\
$\tau$       \> @::=@ \> $rate$                                  \> (loop size of some rate) \\
             \> $~|$  \> @external@                              \> (external and unfusible) \\
\end{tabbing}

@@

$$\begin{array}{lcrcl}
&&\constr{\_} & :: & binds \rightarrow C \\
\\
% MMMMMMMMMMMMMMMMMM \= MM \= \kill
% $\constr{\lambda scalars~arrays \to @let @binds@ in @x}$ \> $=$ \> $\forall_{a \in arrays} k_a. \constr{binds}$ \\

\constr{o &=& @fold @f@ @n}       &  = &  \hole \\
\constr{o &=& @map@_n~f~ns}       &  = &  \forall k_o.\ \bigwedge_{n \in ns}\{k_o = k_n\} \wedge \hole \\
\constr{o &=& @filter@~f~n}       &  = &  \exists k_o \le k_n.\ \hole \\
\constr{o &=& @gather@~i~d}       &  = &  \forall k_o.\ k_o = k_i \wedge \hole \\
\constr{o &=& @cross@~a~b}        &  = &  \forall k_o.\ k_o = k_a \times k_b \wedge \hole \\
\constr{outs &=& @external@~ins}  &  = &  \exists_{o \in outs} k_o.\ \hole \\
\\
&&\constr{b;~bs}  &  = &  \fillhole{\constr{b}}{\constr{bs}}       \\
&&\constr{nil}    &  = &  true                          \\
\end{array}$$

\begin{tabbing}
MM \= MM \= MMMMMMMMM \= MMMM \= MM \= \kill
$\tau$  \>$::$\> $binds \rightarrow name \rightarrow \tau$ \\
\\
$\tau_{bs,o}$    
            \> $|$ \> $o = @fold@~f~n$      \> $\in bs$ \> $=$ \> $k_n$ \\
            \> $|$ \> $o = @map@_n~f~ns$    \> $\in bs$ \> $=$ \> $k_o$ \\
            \> $|$ \> $o = @filter@~f~n$    \> $\in bs$ \> $=$ \> $k_n$ \\
            \> $|$ \> $o = @gather@~i~d$    \> $\in bs$ \> $=$ \> $k_i$ \\
            \> $|$ \> $o = @cross@~a~b$     \> $\in bs$ \> $=$ \> $k_a \times k_b$ \\
            \> $|$ \> $o = @external@~ins$  \> $\in bs$ \> $=$ \> $@external@$ \\
\end{tabbing}

@@
